---
title: "CNQP3"
output: pdf_document
date: "2023-01-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = "")
```

# Load packages

```{r}
library(tidyverse)
library(scales)
library(performance)
library(stargazer)
options(scipen = 999, digits = 2)
```


# Read dataset

```{r}
brazil <- read.csv("brazil.csv")
```


# Question 1

## 1a.

```{r}
sum(is.na(brazil$council.age))
```
* The author have no data on the age of the health council for 99 of the municipalities.


## 1b

```{r, fig.cap="Boxplot of the health council age"}

ggplot(data = brazil, aes(x = 1, y = council.age)) +
  geom_boxplot(outlier.fill = "red", outlier.colour = "red") +
  labs(title = "Boxplot of the health council age",
       y = "Age",
       x = "") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title =element_text(size = 14) ) 
  
```

* The boxplot of the health council age above implies a median of 12 and 1st quartile and 3rd quartile values of 8 and 14 respectively.

* Two noticeable outliers were observed in the boxplot. The outliers are indicated with a red filled color.

## 1c.

```{r}
summary(brazil$corruption)
```
* The mean and median of the `corruption` variable was observed to be 19 and 17 respectively.

* The mean is somewhat higher than the median as it would take into consideration all the values in the sample. As a result, the mean can be easily affected by an outlier. On the other hand, the median is a robust statistic as it's not easily affected by outliers.

* The estimate of the median implies that on the average, the corruption index of the municipalities that the author considered can be taken to be 17.


# Question 2

## 2a.

```{r, fig.cap= "Simple linear regression"}
model1 <- lm(corruption ~ council.age, data = brazil)
stargazer::stargazer(model1, type = "text")
```

## 2b.

* From the output of the simple linear regression above, it could be deduced that both the intercept and the `council.age` variates are significant at 5% level of significant.

* The estimate of the `council.age` implies that there would be on the average a 0.340 reduction in the corruption index as the age of the health council increases by a unit.

* On the other hand, the corruption index for a health council with age of zero is expected to be 23.0. This is the estimate of the intercept.

* The model is significant at 5% level of significant as the p-value of the F-statistic is less than 0.05.


## 2c.

We can interpret the regression coefficient as the average effect of council age on corruption under the following assumptions:

* Linear relationship: The relationship between `corruption` and `council.age` is expected to be linear.

* Independence: Observations are independent of each other

* Homoscedasticity: The variance of the residual is the same for any value of `council.age`.

* Normality: For any fixed value of corruption, council.age is normally distributed.


# Question 3

## 3a.

```{r, fig.cap="Multiple and simple linear models"}
model2 <- lm(corruption ~ council.age + margin + reelected + poverty, data = brazil)

stargazer(model1, model2, type = "text")

```

## 3b.

* The estimated coefficient for margin in the model above is 0.050. This implies an increase in the corruption index for any elected major with a wide margin over the runner-up candidate in the previous election. 

* Thus, the wider the margin between the elected major and the runner-up candidate, the higher the average corruption index.

## 3c.

* The model fit for the multiple linear regression seems to be a better fit than the simple linear regression as the adjusted $R^2$ for the multiple linear regression (0.033 0r 3.3%) is significantly higher than the simple linear regression (0.005 or 0.5%). Both models are significant at 5% level of significant.


## 3d.

```{r}
test_df <- data.frame(council.age = c(10),
                      reelected = c(1),
                      margin = c(12),
                      poverty = c(50))

predict(model2, newdata = test_df)
```

* The predicted corruption index score for a municipality health council that is 10 years old, that has a re-elected Major, where the Major won the last election by 12 percentage points, and where the poverty level is 50 is **19**.


# Question 4

## 4a.

```{r, fig.cap="Multiple linear model with interaction"}
model3 <- lm(corruption ~ council.age + margin + reelected + poverty + council.age*reelected, data = brazil)

stargazer(model2, model3, type = "text")
```

## 4b.

* The estimated coefficient for margin in the model above is 0.054. This implies an increase in the corruption index for any elected major with a wide margin over the runner-up candidate in the previous election. 

* Thus, the wider the margin between the elected major and the runner-up candidate, the higher the average corruption index.


## 4c.

```{r}
confint(model3, "poverty")
```

* For the multiple model with interaction, the 95% confidence interval for the estimate of `poverty` is obtained to be (0.095, 0.21).

## 4d.

```{r}
cor.test(brazil$corruption, brazil$council.age)
```

* From the Pearson product-moment correlation above, it could be deduced that there exist a low negative linear relationship between `corruption` and `council.age`.

## 4e.

```{r}
test_df2 <- data.frame(council.age = c(0,20),
                      reelected = c(1,0),
                      margin = c(10, 10),
                      poverty = c(50, 50))

predict(model3, newdata = test_df2)

plot(predict(model3, newdata = test_df2), ylab = "Predicted Corruption Index", xlab = "")
```

* The plot above also shows that as the `council.age` increases, the corruption index decreases. This is evident as the corruption index for `council.age == 0` is 24 and the corruption index for `council.age == 20` is 19.




