---
title: "Transmuted New Weibull-Pareto Distribution"
author: "Ayodeji Ayodele Theophilus"
date: "4/16/2022"
output: 
  word_document:
    toc: yes
    fig_width: 9
    fig_height: 4.5
  pdf_document:
    toc: yes
  html_document: 
    toc: yes
    toc_float: yes
    anchor_sections: yes
    code_folding: show
    code_download: yes
    highlight: kate
    theme: readable
    number_sections: yes
    fig_width: 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Libraries

```{r}
library(maxLik) 
library(rnoaa)
library(dplyr)
library(ggplot2)
library(lubridate)
options(scipen = 999, digits = 3)
```



# Density Plot

```{r fig.cap="TNWP density plot using different values of the paramters"}

## PDF #=============================================

x <- seq(0, 3, by = 0.0001)
beta <- 1.5
delta <- 0.5
theta <- 0.5
lambda <- 0.0

tnwp.pdf <- function(x, beta=1.5, delta=0.5, 
                     theta=0.5, lambda=0.0){
  (beta*delta/theta) * ((x/theta)^(beta-1))*
    exp(-(delta*(x/theta)^beta)) *(1-lambda + 2*lambda*exp(-(delta*(x/theta)^ beta)))
}


 # par(mfrow = c(2,2), cex.axis = 0.5, cex.lab = 0.5,
 #     mar = c(3.5, 3.5, 3.5, 3.5) - 2.5)

plot(x, tnwp.pdf(x,1.5,0.5,0.5,0.0), type = "l", xlab = "x", 
     ylab = "density", ylim = c(0, 1.7), 
     main = "TNWP density plot using different values of the paramters",
     cex.axis = 0.7, 
     cex.lab = 0.7,
     cex.main = 0.8,
     ann = TRUE)

lines(x, tnwp.pdf(x,1.5,0.5,0.5,0.2), col = 2)
lines(x, tnwp.pdf(x,1.5,0.5,0.5,0.4), col = 3)
lines(x, tnwp.pdf(x,1.5,0.5,0.5,0.6), col = 4)
lines(x, tnwp.pdf(x,1.5,0.5,0.5,0.8), col = 5)
lines(x, tnwp.pdf(x,1.5,0.5,0.5,1.0), col = 6)
legend("topright", cex=0.7, 
       c(expression(paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=0.0"),
                    paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=0.2"),
                    paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=0.4"),
                    paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=0.6"),
                    paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=0.8"),
                    paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=1.0"))),
       col = c(1,2,3,4,5,6),
       lty = c(1,1,1,1,1,1), 
       lwd = c(2,2,2,2,2,2), 
       horiz = FALSE,
       bg = "grey96")


####========== 2nd plot


plot(x, tnwp.pdf(x,1.5,0.5,0.5,0.0), type = "l", xlab = "x", 
     ylab = "density", ylim = c(0, 1.5),
     main = "TNWP density plot using different values of the paramters",
     cex.axis = 0.7, 
     cex.lab = 0.7,
     cex.main = 0.8,
     ann = TRUE)

lines(x, tnwp.pdf(x,1.5,0.5,0.5,-0.2), col = 2)
lines(x, tnwp.pdf(x,1.5,0.5,0.5,-0.4), col = 3)
lines(x, tnwp.pdf(x,1.5,0.5,0.5,-0.6), col = 4)
lines(x, tnwp.pdf(x,1.5,0.5,0.5,-0.8), col = 5)
lines(x, tnwp.pdf(x,1.5,0.5,0.5,-1.0), col = 6)
legend("topright", cex=0.7,
       c(expression(paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=0.0"),
                    paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=-0.2"),
                    paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=-0.4"),
                    paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=-0.6"),
                    paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=-0.8"),
                    paste(beta, "=1.5, ",
                          delta, "=0.5, ",
                          theta, "=0.5, ",
                          lambda, "=-1.0"))),
       col = c(1,2,3,4,5,6),
       lty = c(1,1,1,1,1,1), 
       lwd = c(2,2,2,2,2,2), 
       horiz = F,
       bg = "grey96")


####========== 3rd plot


plot(x, tnwp.pdf(x,3.0,5.0,2.0,0.0), type = "l", xlab = "x", 
     ylab = "density", 
     ylim = c(0, 1.4),
     main = "TNWP density plot using different values of the paramters",
     cex.axis = 0.7, 
     cex.lab = 0.7,
     cex.main = 0.8,
     ann = TRUE)

lines(x, tnwp.pdf(x,3.0,5.0,2.0,0.2), col = 2)
lines(x, tnwp.pdf(x,3.0,5.0,2.0,0.4), col = 3)
lines(x, tnwp.pdf(x,3.0,5.0,2.0,0.6), col = 4)
lines(x, tnwp.pdf(x,3.0,5.0,2.0,0.8), col = 5)
lines(x, tnwp.pdf(x,3.0,5.0,2.0,1.0), col = 6)
legend("topright",cex=0.7, 
       c(expression(paste(beta, "=3.0, ",
                          delta, "=5.0, ",
                          theta, "=2.0, ",
                          lambda, "=0.0"),
                    paste(beta, "=3.0, ",
                          delta, "=5.0, ",
                          theta, "=2.0, ",
                          lambda, "=0.2"),
                    paste(beta, "=3.0, ",
                          delta, "=5.0, ",
                          theta, "=2.0, ",
                          lambda, "=0.4"),
                    paste(beta, "=3.0, ",
                          delta, "=5.0, ",
                          theta, "=2.0, ",
                          lambda, "=0.6"),
                    paste(beta, "=3.0, ",
                          delta, "=5.0, ",
                          theta, "=2.0, ",
                          lambda, "=0.8"),
                    paste(beta, "=3.0, ",
                          delta, "=5.0, ",
                          theta, "=2.0, ",
                          lambda, "=1.0"))),
       col = c(1,2,3,4,5,6),
       lty = c(1,1,1,1,1,1), 
       lwd = c(2,2,2,2,2,2), 
       horiz = F,
       bg = "grey96")

####========== 4th plot


plot(x, tnwp.pdf(x,0.4,0.5,0.0,-1.0), type = "l", 
     xlab = "x", 
     ylab = "density", 
     ylim = c(0, 1.8),
     main = "TNWP density plot using different values of the paramters",
     cex.axis = 0.7, 
     cex.lab = 0.7,
     cex.main = 0.8,
     ann = TRUE)

lines(x, tnwp.pdf(x,0.9,1.2,1.1,0.6), col = 2)
lines(x, tnwp.pdf(x,4.6,2.9,2.8,-0.5), col = 3)
lines(x, tnwp.pdf(x,4.0,1.9,1.5,0.2), col = 4)
lines(x, tnwp.pdf(x,5.5,5.0,6.2,1.0), col = 5)
lines(x, tnwp.pdf(x,5.9,6.2,7.5,1.0), col = 6)
legend("topright",cex=0.7, 
       c(expression(paste(beta, "=0.4, ",
                          delta, "=0.5, ",
                          theta, "=0.0, ",
                          lambda, "=-1.0"),
                    paste(beta, "=0.9, ",
                          delta, "=1.2, ",
                          theta, "=1.1, ",
                          lambda, "=0.6"),
                    paste(beta, "=4.6, ",
                          delta, "=2.9, ",
                          theta, "=2.8, ",
                          lambda, "=-0.5"),
                    paste(beta, "=4.0, ",
                          delta, "=1.9, ",
                          theta, "=1.5, ",
                          lambda, "=0.2"),
                    paste(beta, "=5.5, ",
                          delta, "=5.0, ",
                          theta, "=6.2, ",
                          lambda, "=1.0"),
                    paste(beta, "=5.9, ",
                          delta, "=6.2, ",
                          theta, "=7.5, ",
                          lambda, "=1.0"))),
       col = c(1,2,3,4,5,6),
       lty = c(1,1,1,1,1,1), 
       lwd = c(2,2,2,2,2,2), 
       horiz = F,
       bg = "grey96")
                                                
```


# CDF Plot


```{r fig.cap="CDF of Transmuted Weibull Pareto Distribution"}
## ==== R-code for the CDF plot =========================

x <- seq(0,10,0.025)

cd <- function(x,beta,delta,theta,lambda){
  
  (1-exp(-(delta*(x/theta)^beta)))*
    (1+lambda*(exp(-(delta*(x/theta)^beta))))
  
  }


c1 <- cd(x,beta=1,delta=2,theta=1,lambda=1) 

plot(x,c1, main="CDF of Transmuted Weibull Pareto Distribution",
     type="n", xlab = "x",ylab = "F(x)", 
     cex.main = 0.8, 
     cex.lab = 0.7,
     cex.axis = 0.7) 

lines(x, cd(x,beta=1,delta=2,theta=1,lambda=1), 
      col="red",lwd=2,lty=1) 

lines(x, cd(x,beta=2,delta=2,theta=1,lambda=0),
      col="blue",lwd=2,lty=1) 

lines(x, cd(x,beta=3,delta=2,theta=1.5,lambda=0.5),
      col="green",lwd=2,lty=1) 

lines(x, cd(x,beta=4,delta=2,theta=1.5,lambda=1),
      col="black",lwd=2,lty=1) 


legend("bottomright",cex=0.7, 
       c(expression(paste(beta, "=1, ",
                          delta, "=2, ",
                          theta, "=1, ",
                          lambda, "=1"),
                    paste(beta, "=2, ",
                          delta, "=2, ",
                          theta, "=1, ",
                          lambda, "=0"),
                    paste(beta, "=3, ",
                          delta, "=2, ",
                          theta, "=1.5, ",
                          lambda, "=0.5"),
                    paste(beta, "=4, ",
                          delta, "=2, ",
                          theta, "=1.5, ",
                          lambda, "=1"))),
       horiz=F, 
       lty=c(1,1,1,1),
       lwd=c(2,2,2,2), 
       col=c("red","blue","green","black"), 
       bg="grey96")
                                                  
```

# Survival Plot


```{r fig.cap="Survival Plot of Transmuted Weibull Pareto Distribution"}
## ====R-code for the Survival Plot ======

x <- seq(0,10,0.025) 

h <- function(x, beta, delta, theta, lambda){
  
  (1-((1-exp(-(delta*(x/theta)^beta)))*
        (1+(lambda*(exp(-(delta*(x/theta)^beta)))))))
  
} 

x1 <- h(x,beta=1,delta=2,theta=1,lambda=1) 

x2 <- h(x,beta=2,delta=2,theta=1,lambda=0) 

x3 <- h(x,beta=3,delta=2,theta=1.5,lambda=0.5) 

x4 <- h(x,beta=4,delta=2,theta=1.5,lambda=1)

plot(x, x1, 
     main = "Survival Plot of Transmuted Weibull Pareto Dist.",
     ylim = c(0,1),type = "n", xlab = "x",ylab = "S(x)",
     cex.main = 0.8, cex.lab = 0.7,
    cex.axis = 0.7)

lines(x,x1,col="red",lwd=2,lty=1) 

lines(x,x2,col="blue",lwd=2,lty=1) 

lines(x,x3,col="black",lwd=2,lty=1) 

lines(x,x4,col="green",lwd=2,lty=1) 


legend("topright",cex=0.7, 
       c(expression(paste(beta, "=1, ",
                          delta, "=2, ",
                          theta, "=1, ",
                          lambda, "=1"),
                    paste(beta, "=2, ",
                          delta, "=2, ",
                          theta, "=1, ",
                          lambda, "=0"),
                    paste(beta, "=3, ",
                          delta, "=2, ",
                          theta, "=1.5, ",
                          lambda, "=0.5"),
                    paste(beta, "=4, ",
                          delta, "=2, ",
                          theta, "=1.5, ",
                          lambda, "=1"))),
       horiz=F, 
       lty=c(1,1,1,1),
       lwd=c(2,2,2,2), 
       col=c("red","blue","black","green"), 
       bg="grey96")

```


# Hazard Plot

```{r}
## ==== R-code for the Hazard Plot 

x <- seq(0,10,0.025) 

ha <- function(x,beta,delta,theta,lambda){
  
  ((beta*delta/theta)*((x/theta)^(beta-1))*
     (1-lambda+2*lambda*exp(-(delta*(x/theta)^beta))))/
    (lambda*exp(-(delta*(x/theta)^beta))+1-lambda)
  
  } 

x1 <- ha(x,beta=0.5,delta=3,theta=2,lambda=-1) 

x2 <- ha(x,beta=1,delta=3,theta=2,lambda=0) 

x3 <- ha(x,beta=1.5,delta=3,theta=2,lambda=0.5) 

x4 <- ha(x,beta=2,delta=3,theta=2,lambda=1) 

plot(x,x1, main="Hazard Plot of Transmuted Weibull Pareto Dist.",type="n",xlab = "x",ylab = "H(x)",cex.main = 0.8, 
     cex.lab = 0.7,
     cex.axis = 0.7) 

lines(x,x1,col="red",lwd=2,lty=1)

lines(x,x2,col="blue",lwd=2,lty=1)

lines(x,x3,col="green",lwd=2,lty=1) 

lines(x,x4,col="black",lwd=2,lty=1) 

legend("topright",cex=0.7, 
       c(expression(paste(beta, "=0.5, ",
                          delta, "=3, ",
                          theta, "=2, ",
                          lambda, "=-1"),
                    paste(beta, "=1, ",
                          delta, "=3, ",
                          theta, "=2, ",
                          lambda, "=0"),
                    paste(beta, "=1.5, ",
                          delta, "=3, ",
                          theta, "=2, ",
                          lambda, "=0.5"),
                    paste(beta, "=2, ",
                          delta, "=3, ",
                          theta, "=2, ",
                          lambda, "=1"))),
       horiz=F, 
       lty=c(1,1,1,1),
       lwd=c(2,2,2,2), 
       col=c("red","blue","green","black"), 
       bg="grey96")
```


# Maximum Likelihood Estimation

<!-- ## Public data via `rnoaa` R package -->

<!-- ### Precipitation Data -->

<!-- * The precipitation data is taken from the State of **Florida** station in the United States of America via the `rnoaa` R package. -->

<!-- ```{r comment=""} -->

<!-- # Precipitation -->

<!-- prcp <- ncdc(datasetid='GHCND', stationid='GHCND:USW00014895', datatypeid='PRCP', startdate = '2021-01-01', enddate = '2021-12-31', limit=500, add_units = TRUE, token = Sys.getenv("rnoaa_key")) -->


<!-- prcp_df <- prcp$data -->
<!-- ``` -->


<!-- ```{r fig.width=9} -->
<!-- prcp_df %>%  -->
<!--   mutate(date = ymd_hms(date)) %>%  -->
<!--   ggplot(aes(x = date, y = value)) + -->
<!--   geom_line() + -->
<!--   geom_smooth() + -->
<!--   labs(y = "Precipitation (mm)", -->
<!--        title = "2021 Precipitation (mm) from Florida Station, USA", -->
<!--        x = "") + -->
<!--   scale_x_datetime(date_breaks = "1 month",  -->
<!--                date_labels = "%b-%y") + -->
<!--   theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust = 0.9)) -->
<!-- ``` -->


## Transmuted New Weibull-Pareto

The transmuted new weibull-pareto density function is given below as:

$$f(x) = \frac{\beta~\delta}{\theta}~(\frac{x}{\theta})^{\beta -1}~e^{-\delta}(\frac{x}{\theta})^\beta~(1-\lambda + 2\lambda~e^{-\delta}(\frac{x}{\theta})^\beta)$$

The log likelihood is given below as:

$$loglik = nlog(\beta) + nlog(\delta) -nlog(\theta)-\delta~\sum_{i=1}^n~ (\frac{x_i}{\theta})^\theta + (\beta-1)\sum_{i=1}^n~log(\frac{x}{\theta}) + \sum_{i=1}^n(log(1-\lambda + 2\lambda~e^{-\delta}(\frac{x_i}{\theta})^\beta)$$

***

#### Data

* The dataset used for the analysis was randomly selected from the weibull distribution using the `rweibull` function in R. 

```{r comment=""}
## ===R-code for the Maximum likelihood estimation =====

#x <- prcp_df$value
set.seed(123)
x <- rweibull(120, 2)

loglik <- function(a) {
  
  beta <- a[1]
  delta <- a[2]
  theta <- a[3]
  lambda <- a[4]
  
log_lik <-   n*log(a[1])+n*log(a[2])-n*log(a[3])-(a[2])*
  sum(x/a[3])^a[1]+(a[1]-1)*sum(log(x/a[3]))+
  sum(log(1 - a[4] + 2 * a[4]*exp(-(a[2]*(x/a[3])^a[1]))))
  
  } 

n <- length(x) 

mle_estimate <- maxLik(loglik, 
                       start = c(beta = 3.421, 
                                delta = 8.6231,
                                theta = 2.25,
                                lambda = -0.4608),
                       control=list(printLevel=2)) 

summary(mle_estimate)
AIC(mle_estimate)
```


## Weibull 2-parameter distribution

Let $$f_{\lambda},_{\beta}(x) = \frac{\beta}{\lambda}(\frac{x}{\lambda})^{\beta -1}e^{-(\frac{x}{\beta})^\beta} , ~ x \ge 0$$

The loglikelihood function becomes:

$$loglik = nlog\beta - n\beta~log\lambda - \sum_{i=1}^n~(\frac{x_i}{\lambda})^\beta + (\beta -1)\sum_{i = 1}^n log~x_i$$


```{r comment=""}
weibull_loglik <- function(a) {
  
  beta <- a[1]
  lambda <- a[2]
  
log_lik <-  n*log(a[1]) - n*a[1]*log(a[2]) - (sum(x/a[2])^a[1]) + (a[1] - 1)* sum(log(x)) 
  
  } 

n <- length(x) 

mle_weibull <- maxLik(weibull_loglik, 
                       start = c(beta = 3.421, 
                                lambda = 0.5),
                       control=list(printLevel=2)) 

summary(mle_weibull)

AIC(mle_weibull)
```



# Citation

- R Core Team (2021). R: A language and environment for
  statistical computing. R Foundation for Statistical
  Computing, Vienna, Austria.
  https://www.R-project.org/.
  

- Arne Henningsen and Ott Toomet (2011). maxLik: A package
  for maximum likelihood estimation in R. Computational
  Statistics 26(3), 443-458. DOI
  10.1007/s00180-010-0217-1.
  
- Scott Chamberlain (2021). rnoaa: 'NOAA' Weather Data
  from R. R package version 1.3.8.
  https://CRAN.R-project.org/package=rnoaa
  

