---
title: "Assessment 2"
author: "Halimah"
date: "`r Sys.Date()`"
output: 
  word_document: 
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
```



```{r packages}
library(tidyverse)
library(DescTools)
library(mice)
library(stargazer)
library(car)
options(scipen = 999, digits = 2)
```


## Question 2

```{r}
DiabetesData_Q2 <- read_rds("DiabetesData_Assignment_Q2.RData")
```


### (A) Data Summary

```{r}
DiabetesData_Q2 %>% 
  select(-ID) %>% 
  Desc(.)
```

* The dataset contains 11742 observations with 11 variables. Only 5851(49.8%) complete cases were found in the dataset. 

* The `Gender` variable has no missing value. On the other hand, the `BMI` variable has the most missing values with 3936 (33.5%) missing values.


### (B) Missing Values Imputation

```{r message=FALSE}
set.seed(3333)

new_q2_df <- mice(DiabetesData_Q2, method = "rf") %>% 
  complete()
```

## Question 3

```{r}
DiabetesData_Q3 <- read_rds("DiabetesData_Assignment_Q3.RData") %>% complete()
```


### (A) Fitting a model

**Hypothesis**

* Null hypothesis: $\beta_{MetforminTreatment} = 0$
* Alternative hypothesis:  $\beta_{MetforminTreatment} \neq 0$

```{r}
model1 <- lm(HbA1c ~ MetforminTreatment, data = DiabetesData_Q3)

stargazer(model1, type = "text")
```

**Hypothesis Testing**

```{r}
linearHypothesis(model1, "MetforminTreatment=0")
```


### (B) Model Discussion


* The linear hypothesis test result suggests that the estimate of `MetforminTreatment` is significantly different from zero. This implies that `MetforminTreatment` has a causal effect on `HbA1c`. 
* For the missing value imputation, it is assume that the missing data are missing at Random (MAR), which mean that the probability that a value is missing depends only on observed value and can be predicted using them.




